---
title: "Effects of MPA designation and fishing pressure on the California spiny lobster (_Panulirus interruptus_)"
author: "Joanna Tang, Michelle Lee, & Rachel Behm"

date: "19 November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# INTRODUCTION

The California spiny lobster (_Panulirus interruptus_) is one of the most commercially and economically important lobsters in their native range off the American West Coast (citation). The California Department of Fish and Wildlife (CDFW) estimate that nearly 600,000 individual spiny lobsters were harvested (both recreational and commercial) in the first half of the 2008-2009 season (citation). This number is, on average, half the total harvest from Mexico (citation). Given their economic importance, there are heavy restrictions on the fish season by both the CDFW and the Mexican government, with set harvest season dates, catch limits, and size restrictions (greater than 82.5-82.6 mm) (citation). Moreover, in 2012, Marine Protected Area (MPA) sites were established off the coast of California--excluding fishing activity for the protection of natural and cultural resources. These restrictions insure the sustainability of the fishery and the health of the kelp forest ecosystem in which they live and provide a unique opportunity to understand the direct impacts on spiny lobster populations.
  
While the California spiny lobster remains an important fishery, the species also plays an integral part in the marine communities of the giant kelp forest. _P. interruptus_ is not only an important food source for many predatory species such as the California sheephead, giant sea bass, cabezone, and several species of shark, but it is also a vital predator of sea urchins, clams, muscles, and worms. Given their important role in maintaining ecosystem balance in the kelp forests off the Pacific coast, understanding the human impacts on this species will allow us to better predict the health of these kelp forest ecosystems.
  
MPAs, like those off the coast of California, have been established across the United States marine areas in order to protect ecosystems, preserve cultural resources, and sustain fishery populations (citation). Studies have shown that the implementation of these protected areas have consistently increased the biomass of fished species both inside and outside of MPAs, indicating that these sites, overall, have improved the robustness and health of marine organisms (citation). For many fished species, like the spiny lobster, these biomass increases have important implications for the health and reproductive success as species typically reach sexual maturity with at a certain size. Thus, understanding the direct impacts of MPAs on spiny lobster abundance and size in relation to fishing pressure will allow us to foreshadow the future of the species and the ecosystems where they live. 

Here we will compare California spiny lobster catch and size in the Santa Barbara Channel from 2012 to 2017 to better understand how MPAs affect the species and the giant kelp ecosystem. These data were collected by the Santa Barbara Coast Long Term Ecological Research Program (SBC LTER) as an ongoing effort to monitor the impacts of management on marine habitats.

# METHODS

## _Data Collection_

Lobster size, abundance, and fishing pressure of _P. interruptus_ were collected from 2012-2017 at five SBC LTER Sites off the mainland coast of the Santa Barbara Channel: Arroyo Quemado (AQUE), Naples Reef (NAPL), Mohawk Reef (MOHK), Isla Vista (IVEE), and Carpinteria (CARP). Two of the sites (NAPL and IVEE) are located in or near CDFW MPAs, while the other three (AQUE, MOHK, and CARP) are not. Abundance and size data were collected annually by divers in the late summer before the start of the fishing season. Fishing pressure was determined by counting the number of commerical trap floats. Data were collected every two to four weeks during the lobster fishing season (October to March) (data citation). 

## _Statistical Analyses_

All statistical analyses and figures were created in R (version 3.4.2).

_Lobster size across sites_



_The affect of MPA status_

_Proportion of "legal" lobsters across sites_
To compare the propotion of legal lobster catch sizes across sites, we used a chi-square test 

```{r include = FALSE}
# Outline

# 0. Load packages and read in data
# 1. Create area graph of lobster abundance overlaid with line graph of fishing pressre (5 graphs, 1 per site)
# 2a. Column graph of lobster sizes (1 graph of 5 side-by-sides, 1 per site)
# 2b. ANOVA of lobster sizes (2017)
# 3a. Column graph of lobster sizes (1 graph of 5 side-by-side 2012-2017, 2 columns for each site)
# 3b. 2-sample t-tests comparing size 2012 vs 2017; ANOVA of lobster sizes (2012)
# 4a. Bar graph of proportions of carapace size (1 graph of 5 side-by-sides, 1 per site)
# 4b. Chi-squared of carapace size across sites in 2017

```

```{r include = FALSE}
# 0. Load packages and read in data

library(tidyverse)
library(knitr)
library(kableExtra)
library(effsize)
library(vcdExtra)
library(car)
library(RColorBrewer)
library(ggsignif)

lobster_size_abundance <- read_csv("lobster_size_abundance.csv")
lobster_traps <- read_csv("lobster_traps.csv")


```

```{r echo = FALSE, message = FALSE}

# 1. Lobster abundance and fishing pressure (5 bar + line graphs)

# Data of abundance by site and year
abundance_by_site <- lobster_size_abundance %>% 
  select(YEAR, SITE, COUNT) %>% 
  group_by(SITE, YEAR) %>% 
  summarize(
    abundance = sum(COUNT)
  )

# Data of fishing pressure by site and year
fishing_pressure <- lobster_traps %>% 
  select(YEAR, MONTH, DATE, SITE, TRAPS) %>% 
  filter(SITE == "AQUE" | SITE == "NAPL" | SITE == "MOHK" | SITE == "IVEE" | SITE == "CARP") %>% 
  filter(TRAPS != "-99999") %>% 
  group_by(SITE, YEAR) %>% 
  summarize(
    total_traps_by_month = sum(TRAPS)
    )

abundance_fishing_pressure <- ggplot(abundance_by_site, aes(x = YEAR, y = abundance, displ, cty)) +
  geom_area(aes(fill = SITE), show.legend = FALSE) +
  geom_line(data = fishing_pressure, aes(x = YEAR, y = total_traps_by_month/2)) +
  facet_wrap(~ SITE, scales = "free_x") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0), sec.axis = sec_axis(~.*2, name = "Number of Traps")) +
  theme_classic() + 
  scale_x_continuous(expand = c(0,0)) +
  labs(x= "Year", title="Lobster abundance and fishing pressure (2012 - 2017)", y= "Lobster Abundance") +
  theme(panel.spacing = unit(2, "lines"))

abundance_fishing_pressure



```

Overall, spiny lobster abundance has increased at all sites from 2012 to 2017. At the non-MPA sites (AQUE, CARP, MOHK) fishing pressure has decreased from 2012 to 2017. All sites besides CARP experienced a decrease in spiny lobster abundance in 2015. Fishing pressure and spiny lobster abundance are inversely proportional, the abundance never increased with fishing pressure. 
At site AQUE, spiny lobster abundance has stayed relatively low from 2012 to 2017, with a slight increase during 2013. Fishing pressure at this site remained relatively high until 2016 where there is a steep drop off. At site CARP, spiny lobster abundance also stayed relatively low until 2015 where it started to drastically increase and continued to do so through 2016. Fishing pressure at this site was high until a massive decrease in 2016. At site MOHK, spiny lobster abundance decreased in 2012, slightly increased in 2013, massively increased in 2014, and then began slightly decreasing from 2015 to 2017. 
At the MPA site IVEE, spiny lobster abundance started low in 2012 then increased through 2014, with a drop in 2015, then a drastic rise through 2016. At the MPA site NAPL, spiny lobster abundance steadily grew from 2012 through 2014, with a drop in 2015, and then stabilized through 2016.

```{r echo = FALSE, message = FALSE}

#2a. Compare mean lobster size by site in 2017: Column graph


# Data frame of lobster size and count in 2017
size_count <- as.data.frame(lobster_size_abundance) %>% 
  filter(YEAR == "2017", SIZE != "-99999") %>% 
  select(SITE, SIZE, COUNT)

size_freq <- expand.dft(size_count, freq = "COUNT") # Transformed data into tidy format so every observation is a row


# Summary table of mean and sd of sizes
size_freq_summary <- size_freq %>% 
  group_by(SITE) %>% 
  summarize(
    mean = mean(SIZE),
    sd = sd(SIZE)
  )

# Side-by-side column graphs of mean sizes in 2017
lobster_size <- ggplot(size_freq_summary, aes(x = SITE, y = mean)) +
  geom_col(aes(fill = SITE), show.legend = FALSE) +
  geom_errorbar(aes(ymin = mean-sd, ymax = mean+sd, width = .4)) +
  geom_signif(stat="identity",
              data = data.frame(x=c(1.875, 2.875),
                                xend = c(5,5),
                                y = c(95, 90),
                                annotation = c("*", "**")),
          aes(x=x,
                  xend=xend,
                  y=y,
                  yend=y,
                  annotation=annotation)) +
  labs(x = "Site", y = "Lobster Carapace Length (mm)") +
  theme_classic()
  
lobster_size

```

```{r include = FALSE}
# 2b. Compare mean lobster size by site in 2017: ANOVA

# ANOVA of lobster size (2017)

# First I will make sure I satisfy assumptions of normality and equal variances

size_hists <- ggplot(size_freq, aes(x = SIZE)) +
  geom_histogram() +
  facet_wrap(~ SITE, scale = "free")
size_hists
# Exploratory histograms -- looks normal

size_qq <- ggplot(size_freq, aes(sample = SIZE)) +
  geom_qq() +
  facet_wrap(~ SITE)
size_qq
# Exploratory qq-plots -- looks normal

size_2017_levene <- leveneTest(SIZE ~ SITE, data = size_freq)
size_2017_levene
# p < .001, reject null that variances are equal...BUT I can still do ANOVA if the largest sample variance is <4x the smallest sample variance
variances <- size_freq %>% 
  group_by(SITE) %>% 
  summarize(variance = var(SIZE)
  )
variances
# 86*4 >205 --> can proceed with ANOVA

# ANOVA
# H0: All means of lobster sizes are equal across sites in 2017
# HA: At least 2 means differ

size_2017_anova <- aov(SIZE ~ SITE, data = size_freq)
summary(size_2017_anova)
# F = .0085 --> at least 2 means differ

# Post-hoc Tukey's HSD
TukeyHSD(size_2017_anova)
# NAPL-CARP p=.023, NAPL-IVEE p-.0037, NAPL-MOHK p=.058
```


```{r echo = FALSE, message = FALSE, warning = FALSE}

# 3a. Changes in lobster size at MPA and non-MPA sites (comparing 2012 vs. 2017): Column graph

# Data of just 2012 and 2017
#lobster_size_abundance$YEAR = as.character(lobster_size_abundance)

size_changes_count <- as.data.frame(lobster_size_abundance) %>% 
  filter(YEAR == "2012" | YEAR == "2017", SIZE != "-99999") %>% 
  select(YEAR, SITE, SIZE, COUNT, SBC_LTER_TRANSECT)
size_changes_freq <- expand.dft(size_changes_count, freq = "COUNT") # Transformed data into tidy format so every observation is a row

# Summary table of  2012 and 2017 mean and sd
size_changes_summary <- size_changes_freq %>% 
  group_by(SITE, YEAR) %>% 
  summarize(
    mean = mean(SIZE),
    sd = sd(SIZE), 
    median = median(SIZE)
  )

# Side-by-side (2012 vs 2017) column graphs for mean sizes at non-MPAs
size_changes_graph <- ggplot(size_changes_summary, aes(x = SITE, y = mean)) +
  geom_col(aes(fill = as.factor(YEAR)), position = "dodge", stat = "identity") +
  geom_errorbar(aes(ymin = mean-sd, ymax = mean+sd, width = .4, fill = as.factor(YEAR)), position = "dodge", stat = "identity") +
  labs(x = "Site", y = "Lobster Carapace Length (mm)", fill = "Year") +
  geom_signif(stat="identity",
              data = data.frame(x=c(2.7, 4.4),
                                xend = c(3.3,3.7),
                                y = c(90, 90),
                                annotation = c("*", "**")),
          aes(x=x,
                  xend=xend,
                  y=y,
                  yend=y,
                  annotation=annotation)) +
  theme_classic() +
  scale_fill_brewer(palette = "Paired")

size_changes_graph


```

```{r include = FALSE}
# 3a. Alternate graphs: Boxplots

# Boxplots for each site (2012 vs 2017)
ivee <- ggplot(filter(size_changes_freq, SITE == "IVEE"), aes(x = YEAR, y = SIZE)) +
  geom_boxplot(aes(group = YEAR, fill = YEAR), show.legend = FALSE) +
  labs(x = "Year", y = "Lobster Carapace Length (mm)", title = "Changes in lobster size at Isla Vista from 2012 vs. 2017") +
  theme_classic()
ivee

iveecol <- ggplot(filter(size_changes_freq, SITE == "IVEE", YEAR == "2012" | YEAR == "2017"), aes(x= YEAR,  y= SIZE)) +
  geom_col(aes(group = YEAR, fill = YEAR), show.legend = FALSE) +
  labs(x = "Year", y = "Lobster Carapace Length (mm)", title = "Changes in lobster size at Isla Vista from 2012 vs. 2017") +
  theme_classic()
iveecol
################################################################################################################################
napl <- ggplot(filter(size_changes_freq, SITE == "NAPL"), aes(x = YEAR, y = SIZE)) +
  geom_boxplot(aes(group = YEAR, fill = YEAR), show.legend = FALSE) +
  labs(x = "Year", y = "Lobster Carapace Length (mm)", title = "Changes in lobster size at Naples from 2012 vs. 2017") +
  theme_classic()
napl

naplcol <- ggplot(filter(size_changes_freq, SITE == "NAPL", YEAR == "2012" | YEAR == "2017"), aes(x= YEAR,  y= SIZE)) +
  geom_col(aes(group = YEAR, fill = YEAR), show.legend = FALSE) +
  labs(x = "Year", y = "Lobster Carapace Length (mm)", title = "Changes in lobster size at Naples from 2012 vs. 2017") +
  theme_classic()
naplcol
###############################################################################################################################
aque <- ggplot(filter(size_changes_freq, SITE == "AQUE"), aes(x = YEAR, y = SIZE)) +
  geom_boxplot(aes(group = YEAR, fill = YEAR), show.legend = FALSE) +
  labs(x = "Year", y = "Lobster Carapace Length (mm)", title = "Changes in lobster size at Arroyo Quemado from 2012 vs. 2017") +
  theme_classic()
aque

aquecol <- ggplot(filter(size_changes_freq, SITE == "AQUE", YEAR == "2012" | YEAR == "2017"), aes(x= YEAR,  y= SIZE)) +
  geom_col(aes(group = YEAR, fill = YEAR), show.legend = FALSE) +
  labs(x = "Year", y = "Lobster Carapace Length (mm)", title = "Changes in lobster size at Arroyo Quemado from 2012 vs. 2017") +
  theme_classic()
aquecol
#################################################################################################################################
mohk <- ggplot(filter(size_changes_freq, SITE == "MOHK"), aes(x = YEAR, y = SIZE)) +
  geom_boxplot(aes(group = YEAR, fill = YEAR), show.legend = FALSE) +
  labs(x = "Year", y = "Lobster Carapace Length (mm)",title = "Changes in lobster size at Mohawk from 2012 vs. 2017") +
  theme_classic()
mohk

mohkcol <- ggplot(filter(size_changes_freq, SITE == "MOHK", YEAR == "2012" | YEAR == "2017"), aes(x= YEAR,  y= SIZE)) +
  geom_col(aes(group = YEAR, fill = YEAR), show.legend = FALSE) +
  labs(x = "Year", y = "Lobster Carapace Length (mm)", title = "Changes in lobster size at Mohawk from 2012 vs. 2017") +
  theme_classic()
mohkcol
##################################################################################################################################
carp <- ggplot(filter(size_changes_freq, SITE == "CARP"), aes(x = YEAR, y = SIZE)) +
  geom_boxplot(aes(group = YEAR, fill = YEAR), show.legend = FALSE) +
  labs(x = "Year", y = "Lobster Carapace Length (mm)",title = "Changes in lobster size at Carpinteria from 2012 vs. 2017") +
  theme_classic()
carp

carpcol <- ggplot(filter(size_changes_freq, SITE == "CARP", YEAR == "2012" | YEAR == "2017"), aes(x= YEAR,  y= SIZE)) +
  geom_col(aes(group = YEAR, fill = YEAR), show.legend = FALSE) +
  labs(x = "Year", y = "Lobster Carapace Length (mm)", title = "Changes in lobster size at Carpinteriafrom 2012 vs. 2017") +
  theme_classic()
carpcol
#################################################################################################################################
allbox<-ggplot(size_changes_freq, aes(x = YEAR, y = SIZE)) +
  geom_boxplot(aes(group = YEAR, fill = YEAR), show.legend = FALSE) + facet_wrap(~SITE)
  labs(x = "Year", y = "Lobster Carapace Length (mm)",title = "Changes in lobster size at Carpinteria from 2012 vs. 2017") +
  theme_classic()
allbox 

allcol<-ggplot(size_changes_freq, aes(x = YEAR, y = SIZE)) +
  geom_col(aes(group = YEAR, fill = YEAR), show.legend = TRUE) + facet_wrap(~SITE)
  labs(x = "Year", y = "Lobster Carapace Length (mm)",title = "Changes in lobster size at Carpinteria from 2012 vs. 2017") +
  theme_classic()
allcol  
```



```{r include = FALSE}

# 3b. Changes in lobster size at MPA and non-MPA sites (comparing 2012 vs. 2017): T-tests and ANOVA

# Two-sample (paired) t-tests of lobster size (2012 vs. 2017 for each site)

size_2012_hists <- ggplot(filter(size_changes_freq, YEAR == "2012"), aes(x = SIZE)) +
  geom_histogram() +
  facet_wrap(~ SITE, scale = "free")
size_2012_hists
# Exploratory histograms -- looks normal except for NAPL and IVEE

size_2017_hists <- ggplot(filter(size_changes_freq, YEAR == "2017"), aes(x = SIZE)) +
  geom_histogram() +
  facet_wrap(~ SITE, scale = "free")
size_2017_hists
# Exploratory histograms -- looks normal

size_2012_qq <- ggplot(filter(size_changes_freq, YEAR == "2012"), aes(sample = SIZE)) +
  geom_qq() +
  facet_wrap(~ SITE)
size_2012_qq
# Exploratoery qq-plots -- looks normal except for NAPL and IVEE

size_2017_qq <- ggplot(filter(size_changes_freq, YEAR == "2017"), aes(sample = SIZE)) +
  geom_qq() +
  facet_wrap(~ SITE)
size_2017_qq
# Exploratory qq-plots -- looks normal


# Two-sample t-tests for all sites (2012 vs 2017)
# H0: Mean lobster size in 2012 is the same as mean lobster size in 2017
# HA: Means differ

ivee_var <- filter(size_changes_freq, SITE == "IVEE") %>% 
 var.test(SIZE ~ YEAR, data = .)
ivee_var # p = .307, variances not equal
ivee_2012 <- filter(size_changes_freq, SITE == "IVEE", YEAR == "2012")
ivee_2017 <- filter(size_changes_freq, SITE == "IVEE", YEAR == "2017")
ivee_t <- t.test(ivee_2012$SIZE, ivee_2017$SIZE)
ivee_t # p = .0361, means DIFFERENT (mean increased)

napl_var <- filter(size_changes_freq, SITE == "NAPL") %>% 
 var.test(SIZE ~ YEAR, data = .)
napl_var # p = .77, variances not equal
napl_2012 <- filter(size_changes_freq, SITE == "NAPL", YEAR == "2012")
napl_2017 <- filter(size_changes_freq, SITE == "NAPL", YEAR == "2017")
naple_t <- t.test(napl_2012$SIZE, napl_2017$SIZE)
naple_t # p = .54, means equal

aque_var <- filter(size_changes_freq, SITE == "AQUE") %>% 
 var.test(SIZE ~ YEAR, data = .)
aque_var # p = .298, variances not equal
aque_2012 <- filter(size_changes_freq, SITE == "AQUE", YEAR == "2012")
aque_2017 <- filter(size_changes_freq, SITE == "AQUE", YEAR == "2017")
aque_t <- t.test(aque_2012$SIZE, aque_2017$SIZE)
aque_t # p = .19, means equal

mohk_var <- filter(size_changes_freq, SITE == "MOHK") %>% 
 var.test(SIZE ~ YEAR, data = .)
mohk_var # p = .15, variances not equal
mohk_2012 <- filter(size_changes_freq, SITE == "MOHK", YEAR == "2012")
mohk_2017 <- filter(size_changes_freq, SITE == "MOHK", YEAR == "2017")
mohk_t <- t.test(mohk_2012$SIZE, mohk_2017$SIZE)
mohk_t # p = .00016, means DIFFERENT (mean decreased)

carp_var <- filter(size_changes_freq, SITE == "CARP") %>% 
 var.test(SIZE ~ YEAR, data = .)
carp_var # p = .204, variances not equal
carp_2012 <- filter(size_changes_freq, SITE == "CARP", YEAR == "2012")
carp_2017 <- filter(size_changes_freq, SITE == "CARP", YEAR == "2017")
carp_t <- t.test(carp_2012$SIZE, carp_2017$SIZE)
carp_t # p = .2211, means equal


# Mann-Whitney U's for MPAs
# H0: Ranks of lobster sizes in 2012 is the same as ranks in 2017
# HA: Ranks are different in 2012 and 2017

ivee_mwu <- wilcox.test(ivee_2012$SIZE, ivee_2017$SIZE)
ivee_mwu # p = .0085, ranks different

napl_mwu <- wilcox.test(napl_2012$SIZE, napl_2017$SIZE)
napl_mwu # p = .439, ranks the same

# ANOVA for 2012

# H0: All means of lobster sizes are equal across sites in 2017
# HA: At least 2 means differ

size_2012_anova <- aov(SIZE ~ SITE, data = filter(size_changes_freq, YEAR == "2012"))
summary(size_2012_anova)
# F = .000999 --> at least 2 means differ

# Post-hoc Tukey's HSD
TukeyHSD(size_2012_anova)
# IVEE-CARP p=.025, MOHK-IVEE p=.00062

```

```{r echo = FALSE, message = FALSE}
# 4a. Proportions of "legal" lobsters at the 5 sites in 2017: Bar graph

# Table of proportions of legal lobsters >82.6mm
size_prop <- size_freq %>% 
  group_by(SITE) %>% 
  summarize(
    legal = sum(SIZE >= 82.6)/sum(SIZE>0),
  )

size_prop_graph <- ggplot(size_prop, aes(x = SITE, y = legal)) +
  geom_col(aes(fill = SITE), show.legend = FALSE) +
  labs(x = "Site", y = "Proportion of Legal Lobsters") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0))
size_prop_graph  

```

```{r include = FALSE}

# 4b. Proportions of "legal" lobsters at the 5 sites in 2017: Chi-squared

# Contingency table
legal_size_counts <- size_freq %>% 
  mutate(legal = case_when(SIZE > 82.6 ~ "Legal", SIZE < 92.6 ~ "Too Small")) %>% 
  count(SITE, legal) %>% 
  spread(legal, n) %>% 
  select(-SITE)
rownames(legal_size_counts) <- c("AQUE", "CARP", "IVEE", "MOHK", "NAPL")

# Table of proportions
legal_size_prop <- prop.table(as.matrix(legal_size_counts), 1)

# Chi-squared
# H0: There is no association between proportion of legal lobsters and site
# HA: There is an association

size_chi <- chisq.test(legal_size_counts)
size_chi
# p = .00099, sig diff

```


# Conclusions

Given that spiny lobsters remain, and will remain, economically important for both California and Mexico, understanding the implications of managed and protected lobster habitats is vital for the future of the species and the fishery.

# References

Caselle, JE, Rassweiler, A, Hamilton, SL, and Warner, RR. 2015. Recovery trajectories of kelp forest animals are rapid yet spatially variable across a network of temperate marine protected areas. Scientific Reports: 14102. 

Reed, D. 2017. SBC LTER: Reef: Abundance, size and fishing effort for California Spiny Lobster (_Panulirus interruptus_), ongoing since 2012. Santa Barbara Coastal Long Term Ecological Research Project. doi:10.6073/pasta/81ce20b29614ec99d85d54907eaa3e8e.

